---
title: "Propensity to Cycle Tool - local results"
author: "Created by the PCT team"
output:
  html_document:
    fig_caption: yes
    highlight: pygments
    theme: null
    toc: yes
---

```{r, include=FALSE}
source("set-up.R")
knitr::opts_chunk$set(message = FALSE)

# Read-in the data
if(!exists("region")) region <- "cambridgeshire"
pct_data <- file.path("..", "pct-data")
pct_bigdata <- file.path("..", "pct-bigdata")
zones = readRDS(file.path(pct_data, region, "z.Rds"))
zones$avslope = zones$avslope * 100
cents = readRDS(file.path(pct_data, region, "c.Rds"))
l = readRDS(file.path(pct_data, region, "l.Rds"))
if(!exists("rf_nat"))
  rf_nat <- readRDS(file.path(pct_bigdata, "rf.Rds"))
rf = readRDS(file.path(pct_data, region, "rf.Rds"))
rq = readRDS(file.path(pct_data, region, "rq.Rds"))
rnet = readRDS(file.path(pct_data, region, "rnet.Rds"))
params = readRDS(file.path(pct_data, region, "params.Rds"))
list2env(x = params, envir = globalenv())
source('shared_build.R')
dfscen <- dplyr::select(l@data, contains("slc"), -contains("co2"), All, olc = Bicycle, dist_fast)
dfsp <- gather(dfscen, key = scenario, value = slc, -dist_fast)
dfsp$scenario <- factor(dfsp$scenario)
dfsp$scenario <-
  factor(dfsp$scenario, levels = levels(dfsp$scenario)[c(1, 3, 2, 4, 5, 6)])
levels(dfsp$scenario)[1] <- c("Total no. commuters")
levels(dfsp$scenario)[6] <- c("Cyclists in Census 2011")

las_in_region <- gIntersects(las_cents, region_shape, byid = T)
las_in_region <- las_in_region[1,]
las_in_region <- las[las_in_region,]
```

## Key information 

This document provides information about the data underlying the Propensity to Cycle Tool (PCT) for `r R.utils::capitalize(region)`. The data was generated on `r build_date` and this document was created on `r Sys.Date()`. The PCT is an open source tool for sustainable transport planning, released under the conditions of the [AGPL Licence](https://www.gnu.org/licenses/agpl-3.0). Both the [pct](https://github.com/npct/pct) and [pct-shiny](https://github.com/npct/pct-shiny) can be modified by others as long as attribution is made to the original.

This version of the PCT uses origin-destination (OD) data on travel to work from the 2011 Census. The dataset reports the number of people travelling by different modes from Middle Super Output Area ([MSOA](https://data.gov.uk/dataset/middle-layer-super-output-areas-2001-to-middle-layer-super-output-areas-2011-to-local-authority)) zones. There were `r prettyNum(sum(zones$All), big.mark = ",", scientific = F)` commuters living in `r R.utils::capitalize(region)` recorded in the 2011 Census. All of these are represented in the zones data produced by the PCT.

In `r R.utils::capitalize(region)` there are `r nrow(l)` between-zone flows that a) start and end in `r R.utils::capitalize(region)`, b) have a straight-line (Euclidean) distance of less than `r mdist`km, and c) contain more than `r mflow` commuters (by any mode, counting commuters in both directions).   These `r nrow(l)` between-zone flows are visualised as **Straight Lines**, **Routes** (fast and quiet) and the **Route Network** on the interactive map, and account for `r round(sum(l$All) / sum(zones$All) * 100)`% of all commuters living in `r R.utils::capitalize(region)`.

Between-zone flows exclude **within-zone travel**, when the zone of origin is the same as the zone of destination. Within-zone travel is represented by red points on the map when the lines are shown, and accounts for `r round(sum(cents$All) / sum(zones$All) * 100)`% of commuters in `r R.utils::capitalize(region)`. The between-zone flows visualised as lines and routes on the map also exclude commuters travelling outside `r R.utils::capitalize(region)` and people with no fixed place of work.

See [Lovelace et al. (2016)](http://arxiv.org/abs/1509.04425) for details of the methods used to estimate the cycling, health and carbon impacts of each scenario, and to visualise results at the area, line, route and route network level.

## The `r R.utils::capitalize(region)` region

<!-- Can we show hilliness by comparison to other areas?  -->

```{r distance-dist, echo=FALSE, fig.cap="", echo=FALSE, message=FALSE, warning=FALSE}
plot(zones, col = "lightgrey")
plot(regions, add = T)
plot(las_in_region, border = "blue", add = T, lwd = 2)
plot(region_orig, lwd = 5, add = T)
plot(region_shape, border = "red", add = T, lwd = 2)
lines(l[l$All > 100,], col = "green")
```

**Figure 1:** Map of the study region (thick black border), selected zones (grey), the administrative zone region (red line) and local authorities (blue line). The straight green lines represents the 100 flows with the greatest number of commuters by any mode.

The map below shows the average hilliness of the between-zone routes used by commuters living in each MSOA.
The average hilliness of the fastest routes in
`r R.utils::capitalize(region)` is
`r round(100 * mean(rf$av_incline, na.rm = T), 2)`%,
compared with a national average of
`r round(100 * mean(rf_nat$av_incline, na.rm = T), 2)`%.


```{r, echo = FALSE, fig.cap=""}
tm_shape(zones) +
  tm_fill("avslope", n = 3, palette = "Oranges", title = "Average gradient\nof fast routes (%)") +
  tm_legend()
```

**Figure 2:** Map showing the average hilliness of routes in zones across the study area.

## Cycling levels in each scenario for `r R.utils::capitalize(region)`

The proportion of people who cycled as their main mode of travel to work baseline (2011 Census) data in `r R.utils::capitalize(region)` was 
`r round(100 * sum(zones$Bicycle) / sum(zones$All) , 1)`%, compared to the national average of 3.2%.
The percentage of cycling expected, based on the distance and hilliness of commutes in the region using the baseline model for England, was
`r round(100 * (sum(zones$govtarget_slc) - sum(zones$Bicycle)) / sum(zones$All) , 1)`%.

<!-- Note from Anna: I strongly think we should not include this table commented out here. I think we should just leave a blank until we can give the table that we want, as described below. This table seems quite unclear to me, e.g. I'm not sure who is included in it – I assume it is only between-zone travel, but if the distances go up to 50 km then this is clearly not just the subset of lines that we describe using in the first paragraph. I also wonder if 50 km is correct – do you instead mean 30 km?  In any case, I think better to have a blank for now than this table


The number of people cycling, walking and driving to work in the 2011 Census, by distance band, is illustrated in Table 1 below.  This table only shows between-zone travel.

```{r, echo=FALSE}
dbands = c(0, 2, 5, 10, 15, 20, 50)
l$`Route distance (km)` = cut(l$dist_fast, dbands)
library(dplyr)
dfdb =
  group_by(l@data, `Route distance (km)`) %>% 
  summarise(All = sum(All),
            Cycle = sum(`Bicycle`),
            Drive = sum(`Car_driver`),
            Walk = sum(`Foot`)
            # ,            Other = All - Cycle - Drive - Walk
              )
# dfdb = dfdb[1:5,] # cut out some rows
dfdb$Other = dfdb$All - rowSums(dfdb[3:ncol(dfdb)])
dfdb$`Route distance (km)` = as.character(dfdb$`Route distance (km)`)
dfdb = rbind(dfdb, c("All commuters <50km", colSums(dfdb[2:ncol(dfdb)])))
htmlTable::htmlTable(dfdb, caption = '<strong>Table 1:</strong> The number of people travelling by distance band and mode in the region in the 2011 Census. Note that ‘(0,2]’ means ‘from more than zero, up to and including 2’, as defined by <a href="https://en.wikipedia.org/wiki/ISO_31-11">ISO 31-11</a>.')
```
-->


<!-- Summary results for the proportion of people walking, driving and using other modes under each scenario are illustrated in the table below. -->

<!--As discussed with Anna - can we please have here a table for the whole region per scenario, using area-based data? This would look like the following
                              % cyclists  % walking % car drivers % all other modes
  Census 2011
  Government Target
  Gender Equity
  Go Dutch
  Ebikes
-->

Figure 3 illustrates the overall number of people in between-zone flows cycling to work under each scenario (note that this does not include within-zone travel).

<!--I think we Ideally should revisit this figure at some point and include within-zone flows, I think it would make the figure much more helpful
-->
```{r, echo=FALSE, warning=FALSE, fig.cap=""}
levels(dfsp$scenario) <- gsub("_slc", "", levels(dfsp$scenario))
levels(dfsp$scenario) <- R.utils::capitalize(levels(dfsp$scenario))
levels(dfsp$scenario)[2:5] = c("Cyclists in Ebikes", "Cyclists in Go Dutch", "Cyclists in Gender Equality", "Cyclists in Government Target")
ggplot(dfsp) +
  geom_freqpoly(aes(dist_fast, weight = slc,
    color = scenario), binwidth = 1) +
  ylab("Number of commuters") +
  xlab("Route distance (km)") +
  scale_color_discrete(name = "Scenario") +
  xlim(c(0,12)) +
  theme_bw()
```

**Figure 3:** Number of between-zone commuters within `r R.utils::capitalize(region)` cycling by scenario and distance.

<!-- For further information on -->

## References

Lovelace, Robin, Anna Goodman, Rachel Aldred, Nikolai Berkoff, Ali Abbas, and James Woodcock. 2016. "The Propensity to Cycle Tool: An Open Source Online System for Sustainable Transport Planning."" ArXiv:1509.04425 [Cs]. http://arxiv.org/abs/1509.04425.




<!-- ## Health and carbon impacts -->

<!-- A modified -->
<!-- version of the 2014 [HEAT tool](http://www.euro.who.int/en/health-topics/environment-and-health/Transport-and-health/publications/2011/health-economic-assessment-tools-heat-for-walking-and-for-cycling.-methodology-and-user-guide.-economic-assessment-of-transport-infrastructure-and-policies.-2014-update) -->
<!-- was used to calculate health impacts, based on the extra physical activity gained through new cycling trips in each scenario, minus the physical activity displaced in any former walking trips that have now been switched to cycling. CO2 impacts are derived from changes in distance driven per scenario. -->

<!-- The table below illustrates the health and carbon impacts of the different scenarios. -->

<!-- please create table below.

Scenario      Comparison                  Mortality Impacts                        Change in MTCO2e/year
                              Change in deaths/yr Value of change (£, millions)
Census 2011  'No Cyclists'
Govt Target 'No Cyclists'
            Census 2011
Gender equity 'No Cyclists'
            Census 2011
Go Dutch 'No Cyclists'
            Census 2011
Ebikes 'No Cyclists'
            Census 2011-->

