---
title: "PCT - local LSOA school results"
author: "Created by the PCT team"
output: 
  html_document:
    fig_caption: yes
    highlight: pygments
    theme: null
    toc: yes
---

```{r, include=FALSE}
options(scipen=999)
knitr::opts_chunk$set(message = FALSE)

# Read-in the data
root <- "../.."
d <- readRDS(file.path(root, path_outputs_regional_R, purpose_private, geography, region, "d.Rds"))
z <- readRDS(file.path(root, path_outputs_regional_R, purpose_private, geography, region, "z.Rds"))

# Import data for Fig 1 + 2 maps
region_shape <- regions_highres[regions_highres$region_name==region,]
las_in_region <- las[las@data$lad11cd %in% region_lad_lookup$lad11cd, ]

# Generate width, height and aspect for region figures
region_shape_ne <- spTransform(region_shape, proj_27700)
region_width <- (region_shape_ne@bbox[1,2] - region_shape_ne@bbox[1,1])/1000
region_height <- (region_shape_ne@bbox[2,2] - region_shape_ne@bbox[2,1])/1000
region_aspect <- region_height/region_width
region_figwidth <- NULL
if (region_aspect<0.3) {region_figwidth <- 8.5}
if (region_aspect>=0.3 & region_aspect<0.5) {region_figwidth <- 8}
if (region_aspect>=0.5 & region_aspect<0.7) {region_figwidth <- 7.5}
if (region_aspect>=0.7 & region_aspect<0.9) {region_figwidth <- 7}
if (region_aspect>=0.9 & region_aspect<1.1) {region_figwidth <- 6}
if (region_aspect>=1.1 & region_aspect<1.5) {region_figwidth <- 5}
if (region_aspect>=1.5 & region_aspect<1.9) {region_figwidth <- 4.5}
if (region_aspect>=1.9 & region_aspect<2.3) {region_figwidth <- 3.5}
if (region_aspect>=2.3) {region_figwidth <- 2.5}


```

### Key information for `r get_pretty_region_name(region)`, school layer, LSOA geography

This document provides information about the data underlying the Propensity to Cycle Tool (PCT) schools layer for `r get_pretty_region_name(region)`. The data were generated on `r region_build_param$build_date` and this document was created on `r Sys.Date()`. The PCT is an open source tool for sustainable transport planning, released under the conditions of the 
<a target="_blank" href="https://www.gnu.org/licenses/agpl-3.0">Affero GPL</a>. The <a target="_blank" href="https://github.com/npct">PCT source code</a> can be modified by others as long as attribution is made to the original.

This layer of the PCT uses origin-destination data on travel to school from the 2011 National School Census (NSC). The interface is currently displaying the number of children travelling by different modes to school from the Lower Super Output Area (LSOA) where they live (LSOAs have an average size 360 children). 

There were `r prettyNum(sum(z$all), big.mark = ",", scientific = F)` school children living in `r get_pretty_region_name(region)` recorded in the 2011 NSC. All of these children are included in the **zones** data that are visualised on the map interface and available for download. The interface pop-ups do not, however, present exact numbers for counts based on 1 or 2 children.

There were `r prettyNum(sum(d$all), big.mark = ",", scientific = F)` school children attending a school in `r get_pretty_region_name(region)` recorded in the 2011 NSC. All of these children are included in the **destinations** data that are visualised on the map interface and available for download.  The interface pop-ups do not, however, present exact numbers for counts based on 1 to 5 children.

Finally, the **Route Network (LSOA, clickable)** is built from all origin-destination fast routes within the region that contain at least 5 children and have a fast route distance <5km (for primary school children)/<10km (for secondary school children). The pop-ups do not present exact numbers for segments containing 1 to 2 cyclists.

For more details of methods please see the User Manual.

<br>
### The `r get_pretty_region_name(region, the=F)` region

Figure 1 shows a map of `r get_pretty_region_name(region)`, including the regional boundary (thick black border), local authority boundaries (blue lines) and LSOA zones in the region (grey with a black border). 

*Figure 1: Map of the `r get_pretty_region_name(region, the=F)` region*
```{r fig1, fig.align = "left", fig.cap="", echo=FALSE, message=FALSE, warning=FALSE}
par(mar=c(0.2, 0.2, 0.2, 0.2) + 0.1)
plot(z, col = "lightgrey")
plot(las_in_region, border = "blue", lwd = 2, add = T)
plot(regions_highres, lwd = 2, add = T)
plot(region_shape, lwd = 5, add = T)
```
The left-hand map in Figure 2 shows the proportion of children in each zone with a fast route distance to school less than 5km.  The average proportion was `r round(mean(z$perc_rf_dist_u5km, na.rm = T))`% across zones in `r get_pretty_region_name(region)`, compared with a national average of `r round(mean(z_all$perc_rf_dist_u5km, na.rm = T))`%. The right-hand map shows the average hilliness of the fastest routes used for travelling to school in each zone.  The average was `r sprintf("%.1f", mean(z$avslope_perc_u5km, na.rm = T))`% across zones in `r get_pretty_region_name(region)`, compared with a national average of `r sprintf("%.1f",  mean(z_all$avslope_perc_u5km, na.rm = T))`%.

*Figure 2: Map showing a) the proportion of children with a fast route distance to school <5km (left) and b) the average hilliness of school trips <5km (right), in LSOA zones across the study area*
```{r fig2, fig.width = (region_figwidth*2), fig.asp = (region_aspect/2), fig.align = "left", fig.cap="", echo = FALSE, message=FALSE, warning=FALSE}
tm_shape(z) +
  tm_fill(col=c("perc_rf_dist_u5km","avslope_perc_u5km"),
              n=c(4,4),
              palette=list("Blues", "Oranges"),
              border.col = "black",
              title=c("% school trips\n<5km", "Average\ngradient (%)")) +
  tm_layout(legend.title.size=1.1, legend.text.size = 1.1, frame = F) +
  tm_facets(ncol = 2, nrow=1, scale.factor=4)

```

<br>

### Cycling levels in each scenario for `r get_pretty_region_name(region)`

At baseline (NSC 2011), the proportion of commuters who cycled to school in `r get_pretty_region_name(region)` was 
`r round(100 * sum(z$bicycle) / sum(z$all) , 1)`%, compared to the national average of 1.8% in England as a whole.  The percentage of cycling expected was
`r round(100 * (sum(z$govtarget_slc) - sum(z$bicycle)) / sum(z$all) , 1)`%, based on the distance and hilliness of school trips in the region using the baseline model for England.

Table 1 summarises the proportion of children cycling, walking, driving and using other modes under each scenario in `r get_pretty_region_name(region)`, based on aggregating across all residents in all the LSOA zones.

*Table 1: Travel to school mode share in the NSC 2011 and in scenarios among `r prettyNum(sum(z$all), big.mark = ",", scientific = F)` children living in `r get_pretty_region_name(region)`* 

| Scenario | % cycling | % walking | % driven by car | % all other modes |
| ------------- |:-------------: |:-------------: |:--------------:| :----------------: |
| NSC 2011 | `r paste(sprintf('%.1f', (b <- round(100 * sum(z$bicycle) / sum(z$all) , 1))), '%')`|`r paste(sprintf('%.1f', (f <- round(100 * sum(z$foot) / sum(z$all) , 1))), '%')`|`r paste(sprintf('%.1f', (cd <- round(100 * sum(z$car) / sum(z$all) , 1))), '%')` | `r paste(sprintf('%.1f', (100 - b - f - cd)), '%')` |
| Government Target | `r paste(sprintf('%.1f', (b <- round(100 * sum(z$govtarget_slc) / sum(z$all) , 1))), '%')` | `r paste(sprintf('%.1f', (f <- round(100 * sum(z$govtarget_slw) / sum(z$all) , 1))), '%')`|`r paste(sprintf('%.1f', (cd <- round(100 * sum(z$govtarget_sld) / sum(z$all) , 1))), '%')` | `r paste(sprintf('%.1f', (100 - b - f - cd)), '%')` |
| Go Dutch | `r paste(sprintf('%.1f', (b <- round(100 * sum(z$dutch_slc) / sum(z$all) , 1))), '%')` | `r paste(sprintf('%.1f', (f <- round(100 * sum(z$dutch_slw) / sum(z$all) , 1))), '%')`|`r paste(sprintf('%.1f', (cd <- round(100 * sum(z$dutch_sld) / sum(z$all) , 1))), '%')` | `r paste(sprintf('%.1f', (100 - b - f - cd)), '%')` |

<br>
